---
title: "Analysis of Productivity Stats in NZ from 1978 - 2022"
output: pdf_document
date: "2024-01-18"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

#Packages
library(ggplot2)
#Read in the data
data <- read.csv("~/R projects/population-stats-in-r/productivity-statistics-1978-2022.csv", header=TRUE)
data <- data[, 1:11]

#Renaming of the data columns
new_data_col_names <- c("Series_reference","Period","Data_value","Status","Units","Magnitude",
                        "Subject","Group","Type","Industry","Measure")
names(data) <- new_data_col_names
non_blank_industries <- data[data$Industry != "",]
```

## Analysis of Productivity Stats in NZ from 1978 - 2022

This is a report to investigate the productivity statistics that were provided by [Stats NZ](https://www.stats.govt.nz/large-datasets/csv-files-for-download/). I've chosen to look through the data to find any meaningful insights as to how the different industries have changed over the 44 years of data.

#Introduction

A glance into the data indicates a large variety of productivity measures contained within. The first rows of the data are as follows:

```{r}
head(data)
```

A summary of the data provided below:
```{r}
summary(data)
```

The structure of the data:

```{r}
str(data)
```

# 2022 Q3 results

The initial year of 1978 was considered the base year with all industries getting a measure of their productivity as 1000. Below is a look at how the industries fared in 2022:

```{r}
prod_data_2022 <- data[data$Period == "2022.03" & data$Measure == "Index" & data$Type == "Output",]
library(ggplot2)

# Will plot the top 5 and bottom 5 industries
ggplot(data = prod_data_2022, aes(x = Data_value, y = Industry)) +
  geom_bar(stat = "identity", fill = "skyblue", color = "black") +
  labs(title = "Productivity of all industries in 2022", x = "Productivity measure", y = "Industry") +
  theme_minimal()







```

Taking a closer look into the top 5 most productive industries - which by the design of the data, have had the highest percentage change since inception of the measure in NZ.

```{r}
ordered_2022_data <- prod_data_2022[order(prod_data_2022$Data_value),]
top_5_industries <- ordered_2022_data[-(1:17),]
ggplot(data = top_5_industries, aes(x = Data_value, y = Industry)) +
  geom_bar(stat = "identity", fill = "skyblue", color = "black") +
  labs(title = "Productivity of top 5 industries in 2022", x = "Productivity measure", y = "Industry") +
  theme_minimal()
```

Now looking into the lowest productive industries:

```{r}
bottom_5_industries <- ordered_2022_data[-(6:22),]
ggplot(data = bottom_5_industries, aes(x = Data_value, y = Industry)) +
  geom_bar(stat = "identity", fill = "skyblue", color = "black") +
  labs(title = "Productivity of bottom 5 industries in 2022", x = "Productivity measure", y = "Industry") +
  theme_minimal()
```


Look to complete report - unsure on what some measures are.

#Results
I should definitely have researched what is involved in productivity statistics and data before beginning on this endeavour. I managed to trim down the dataset by removing empty columns and understanding the data frame supplied. There was only one column on measurements which made it easier to get my head around. From there it was understanding that the initial year was considered a baseline year with all data values given as 1000. 

Information media and communication was by far the most productive industry of all of those measured. This isn't surprising due to the great techonological age that we live in. It is also a service rather than a good which helps with scale of outputs. 

Most of the other industries are service industries which seems logical as they tend to be easier to scale than goods based businesses. This is why it is surprising to see retail trade as one of the top 5 most productive industries in 2022. 

The industries at the bottom aren't too surprising. Most of them are goods related industries which are harder to scale. These industries haven't had large technological innovations in the last 40 years. Education and training is one that is a little surprising. I would have predicted that with the rise of online courses that the productivity would be higher. I'm not certain what the inputs are for training and education and whether it includes public schools, universities, polytechnics or any form of training. 





