---
title: "Analysis of Productivity Stats in NZ from 1978 - 2022"
output: pdf_document
date: "2024-01-18"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

#Packages
library(ggplot2)
#Read in the data
data <- read.csv("~/R projects/population-stats-in-r/productivity-statistics-1978-2022.csv", header=TRUE)
data <- data[, 1:11]

#Renaming of the data columns
new_data_col_names <- c("Series_reference","Period","Data_value","Status","Units","Magnitude",
                        "Subject","Group","Type","Industry","Measure")
names(data) <- new_data_col_names
non_blank_industries <- data[data$Industry != "",]
```

## Analysis of Productivity Stats in NZ from 1978 - 2022

This is a report to investigate the productivity statistics that were provided by [Stats NZ](https://www.stats.govt.nz/large-datasets/csv-files-for-download/). I've chosen to look through the data to find any meaningful insights as to how the different industries have changed over the 44 years of data.

#Introduction

A glance into the data indicates a large variety of productivity measures contained within. The first rows of the data are as follows:

```{r}
head(data)
```

A summary of the data provided below:
```{r}
summary(data)
```

The structure of the data:

```{r}
str(data)
```

The initial year of 1978 was considered the base year with all industries getting a measure of their productivity as 1000. Below is a look at how the industries fared in 2022:

```{r}
prod_data_2022 <- data[data$Period == "2022.03" & data$Measure == "Index" & data$Type == "Total",]
library(ggplot2)

# Will plot the top 5 and bottom 5 industries
ggplot(data = prod_data_2022, aes(x = Data_value, y = Industry)) +
  geom_bar(stat = "identity", fill = "skyblue", color = "black") +
  labs(title = "Productivity of all industries in 2022", x = "Productivity measure", y = "Industry") +
  theme_minimal()







```

Taking a closer look into the top 5 most productive industries - which by the design of the data, have had the highest percentage change since inception of the measure in NZ.

```{r}
ordered_2022_data <- prod_data_2022[order(prod_data_2022$Data_value),]
top_5_industries <- ordered_2022_data[-(1:17),]
ggplot(data = top_5_industries, aes(x = Data_value, y = Industry)) +
  geom_bar(stat = "identity", fill = "skyblue", color = "black") +
  labs(title = "Productivity of top 5 industries in 2022", x = "Productivity measure", y = "Industry") +
  theme_minimal()
```

Now looking into the lowest productive industries:

```{r}
bottom_5_industries <- ordered_2022_data[-(6:22),]
ggplot(data = bottom_5_industries, aes(x = Data_value, y = Industry)) +
  geom_bar(stat = "identity", fill = "skyblue", color = "black") +
  labs(title = "Productivity of bottom 5 industries in 2022", x = "Productivity measure", y = "Industry") +
  theme_minimal()
```
Having a look into the types of productivity in the top 5 industries:
```{r}
top_5_industries_by_name <- top_5_industries[,"Industry"]
types_of_interest <- c("Capital", "Labour", )
all_industry_prod_data_2022 <- data[data$Period == "2022.03" & data$Measure == "Index" & data$Group == "Productivity Indexes - Industry Level (ANZSIC06)",]
top_5_industry_prod_data_2022 <- all_industry_prod_data_2022[all_industry_prod_data_2022$Industry %in%  top_5_industries_by_name,]

ggplot(data = top_5_industry_prod_data_2022, aes(x = Data_value, y = Industry, fill=Type)) +
  geom_bar(stat = "identity", color = "black") +
  labs(title = "Productivity of bottom 5 industries in 2022", x = "Productivity measure", y = "Industry") +
  theme_minimal()
```


