---
title: "Analysis of Productivity Stats in NZ from 1978 - 2022"
output: pdf_document
date: "2024-01-18"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
#Read in the data
data <- read.csv("~/R projects/population-stats-in-r/productivity-statistics-1978-2022.csv", header=TRUE)
data <- data[, 1:11]

#Renaming of the data columns
new_data_col_names <- c("Series_reference","Period","Data_value","Status","Units","Magnitude",
                        "Subject","Group","Type","Industry","Measure")
names(data) <- new_data_col_names
```

## Analysis of Productivity Stats in NZ from 1978 - 2022

This is a report to investigate the productivity statistics that were provided by [Stats NZ](https://www.stats.govt.nz/large-datasets/csv-files-for-download/). I've chosen to look through the data to find any meaningful insights as to how the different industries have changed over the 44 years of data.

#Introduction

A glance into the data indicates a large variety of productivity measures contained within. The first rows of the data are as follows:

```{r}
head(data)
```
